\chapter{Github Repository}
Appendix B includes links to a github repository containing codes described in the thesis. Here is a link to the repository main:


\url{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid}

The following section contains hyperlinks to certain parts of the repository. If the reference in the thesis says "see Appendix B.4" it refers to Matlab file 'Resistor\_model\_8x8\_full.m' which is the fourth file in the repository.

\newlist{continuing}{enumerate}{1}
\setlist[continuing]{label=\arabic*.,resume}

Matlab files:
    \begin{continuing}
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Matlab/Force_program.m}{'Force\_program.m'}
        contains the code presenting sensor readings in a three dimensional plot, fitted to display the values in Force.
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Matlab/Resistor_model_3x3_full.m}{'Resistor\_model\_3x3\_full.m'} contains the code to calculate equations for 3 by 3 resistor model and then executing the solve.
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Matlab/Resistor_model_3x3_with_equation.m}{'Resistor\_model\_3x3\_with\_equations.m'} contains the code to only execute the solve where the equation are already calculated.
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Matlab/Resistor_model_8x8_full.m}{'Resistor\_model\_8x8\_full.m'} contains the code to calculate equations for 8 by 8 resistor model and then executing the solve.
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Matlab/Test_single.m}{'Test\_single.m'} contains a test program used to read and plot voltage, force and resistance as a function of time.
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Matlab/Voltage_program.m}{'Voltage\_program.m'} contains the code presenting sensor readings in a three dimensional plot in voltage.
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Matlab/read_specs.m}{'read\_specs.m'} contains a file containing the function to read from the specification file.
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Matlab/specs.txt}{'specs.txt'} contains the specification file.
    \end{continuing}
    
Arduino IDE files:
    \begin{continuing}
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Arduino/8x8_program.ino}{'8x8\_program.ino'} contains the general code to read data from the modules which can be used together with both the 'Force\_program' and 'Voltage\_program' in Matlab.
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Arduino/Motor_control.ino}{'Motor\_control.ino'} contains the Arduino code to run the motor for the tests.
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Arduino/Test_single.ino}{'Test\_single.ino'} contains the Arduino code matching the 'Test\_single.m' Matlab code, where you manually choose the location of the sensor to be read in the Arduino code.
    \end{continuing}
    
Maple files:
    \begin{continuing}
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Maple/Resistor_model_3x3.mw}{'Resistor\_model\_3x3.mw'} contains the maple document with the calculation to make and execute the 3 by 3 resistor model
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Maple/Resistor_model_3x3.pdf}{'Resistor\_model\_3x3.pdf'} contains the pdf of the maple document 'Resistor\_model\_3x3.mw'
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Maple/Uniform_pressure.mw}{'Uniform\_pressure.mw'} contains the maple document with the equations and formula resistor models where resistors are exerted with uniform pressure
        \item 
        \href{https://github.com/s214413/Force-Displacement-Visualization-through-3D-Printed-Sensor-Grid/blob/main/Maple/Uniform_pressure.pdf}{'Uniform\_pressure.pdf'} contains the pdf of the maple document 'Uniform\_pressure.mw'
    \end{continuing}